---
title: "Learning Strategies for Poker with a Genetic Algorithm"
author: "Raphaël Morsomme"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: TRUE
---

```{r libraries, message=F}
library(tidyverse)
library(gdata)
```

# Introduction
In this script, I use a traditional genetic algorithm (GA) to learn strategies for a fairly simple version of poker. I investigate poker because I love playing this game, and to use a GA because it offers a nice coding challenge and is a suitable method for learning good strategies for this game. A question that I want to answer at the end of the script is if the GA will learn strategies containing an element of bluff.

# Background

## A Simple Version of Poker
We consider a fairly simple version of poker opposing two players: player A and player B. A hand starts with each player paying an *ante* (small predetermined amount) and receiving a single card. Player A then effectuates a *bet* and player B decides to either *fold* or *call* it. If player B folds, then player A wins the hand and recuperates the pot (the two antes). If player B calls player A's bet, then the players show their cards and the one with the highest card wins the hand and recuperates the pots (the two antes and the two bets). If player B calls player A's bet and both players have the same card, they share the pot.

In short, a hand follows **5 steps** in which we determine:

1. **Player A's bet** based on player A's strategy and player A's card.
2. **Player B's action** based on player B's strategy, player B's card and player A's bet.
3. **The size of the pot** based on player A's bet and player B's action.
4. **The winner of the hand** given the cards of the two players and player B's action.
5. **The gain/loss** of each player at the end of the hand.

Before delving into the GA itself, we examine the nature of the strategies, and how to simulate a hand.

## Strategies
Player A's strategy consists of a vector of numeric values indicating the amount to bet for each card (s)he could receive. If we consider a game where player A can effectuate the following bets `r seq(0,10,2)` and the cards are `r 1:4`, the following is a possible strategy for player A
```{r strategy A example, echo=F}
strategy_A <- array(c(0, 10, 2, 8), dim = c(4,1), dimnames = list(1:4, "Strategy A"))
# Strategy for player A
print(strategy_A)
```
which prescribes to bet `0` if (s)he receives a card `1`, `10` (a bluff!) if (s)he receives a `2`, `2` if (s)he receives a `3`, and to bet `8` if (s)he receives a `4`.

Player B's strategy consists of a matrix whose entries indicate the action to realise - to *fold* or to *call* - for each possible combination of her/his card and player A's bet. The following is a possible strategy for player B
```{r strategy B example, echo=F}
strategy_B <- array(c("Fold", "Fold", "Call", "Fold", "Call", "Fold", "Fold", "Fold", "Call", "Fold", "Call","Call", "Fold","Call", "Fold", "Call","Call", "Fold", "Fold", "Fold", "Call", "Fold", "Fold", "Call"), dim=c(6,4, 1), dimnames=list(seq(0, 10, 2), 1:4, "Strategy B"))
# Strategy for Player B
print(strategy_B)
```
which prescribes to `fold` if (s)he receives a card `1` and player A bets `0`, to `call` if (s)he receives a `2` and player A bets `4` and to `call` if (s)he receives a `4` and player A bets `10`.

## Simulating a Hand

### Setup
We assign the value of the ante, the cards we play with and the possible bets for player A to `ante`, `cards` and `bets` respectively.
```{r setup}
ante   <- 2
cards  <- 1:4        
bets   <- seq(from = 0, to = 10, by = 2)

# for convenience
n_card      <- length(cards)
n_bet       <- length(bets)
dim_mat     <- c(n_card, n_card)  # (see matrix-oriented approach)
dimname_mat <- list(cards, cards) # (see matrix-oriented approach)
```

We use `sample()` in `array()` to create random strategies for the two players and assign them to `strategy_A` and `strategy_B`.
```{r Random Strategies}
set.seed(123)
strategy_A <- array(sample(x = bets, size = n_card, replace = T), 
                    dim = c(n_card, 1), dimnames = list(cards, "Strategy A"))
strategy_B <- array(sample(x = c("Call", "Fold"), n_card * n_bet, T),
                    dim=c(n_card, n_bet, 1), dimnames = list(cards, bets, "Strategy B"))
# Random strategy for player A
print(strategy_A)
# Random strategy for player B
print(strategy_B)
```

### The Naive Approach
To simulate a hand, one could be tempted to imitate what would happen if two players sat down around a table to play a hand. Such code would be painfully slow for simulating a large number of hands - which is exactly what the GA requires - as one would need to use loops (slow on `R`) to accomplish this. We therefore need a more efficient approach.

### The Matrix-Oriented Approach
The *matrix-oriented* approach uses matrices and matrix operations (fast on `R`) to simulate a large number of hands at once. We generate five matrices (one for each step of a hand) whose columns and rows respectively correspond to player A's and player B's cards. This way, we can simulate all possible hands between two given strategies at once. This makes the matrix-oriented approach much more efficient than the naive one. The matrices' entries respectively represent (i) player A's bet (ii) player B's action (iii) the pot size (iv) the winner of the hand and (v) player A's gain/loss.

1. **Determining player A's bet**

The matrix `bet_A` represents player A's bet for each possible hand. Since player A's bet only depends on her/his card, each column of `bet_A` contains the same values.
```{r step 1}
bet_A <- array(rep(strategy_A, each = n_card), dim = dim_mat, dimnames = dimname_mat)
# Player A's bet
print(bet_A)
```

2. **Determining player B's action**

The matrix `action_B` represents player B's action. Since `bet_A`'s columns are uniform, `action_B` is simply a re-ordering of `strategy_B`'s columns.
```{r step 2}
action_B <- array(strategy_B[, match(strategy_A, bets), 1], dim = dim_mat, dimnames = dimname_mat)
# Player B's strategy
print(strategy_B)
# Player B's action
print(action_B)
```

3. **Determining the size of the pot**

`pot_size` indicates the size or the pot. If player B folds, then the pot only consists of the two antes.
```{r step 3}
pot_size                     <- 2 * ante + 2 * bet_A 
pot_size[action_B == "Fold"] <- 2 * ante
# Size of the pot
print(pot_size)
```

4. **Determining the winner of hand** 

`win_game` indicates the winner of the hand (`1`: player A wins, `0`: a draw and `-1`: player B wins).
```{r step 4, message=F}
win_game <- array(numeric(n_card * n_card), dim = dim_mat, dimnames = dimname_mat)

upperTriangle(win_game, diag = F) <-  1
lowerTriangle(win_game, diag = F) <- -1

win_game[action_B == "Fold"] <- 1
# Winner of the hand
print(win_game)
```

5. **Determining player A's gain/loss**

`gain_A` represents the amount won by player A. A negative amount indicates that player A loses money. Since this is a zero-sum game between two players, player B's gain/loss is simply the opposite of player A's.
```{r step 5}
gain_A <- pot_size * win_game / 2
# Player A's gain/loss
print(gain_A)
# Player B's gain/loss
print(- gain_A)
```

# The Genetic Algorithm

## Overview
In order to learn good strategies for our two players, we use a GA. Simply put, a GA simulates how a process of *natural selection* iteratively selects from an existing population the fittest individuals and cross over their genes to generate new individuals. A *good* GA produces, after some time, a population of fit individuals. In this script, we design a GA that learns profitable strategies for the two players. The question that we want to answer is whether the strategies that the GA will learn will contain an element of bluff.

In practice, we first consider an initial *population* of strategies for each player; these are the strategies with which we start the GA[^3]. We then make each strategy of player A's population play against each strategy of player B's population. Based on the results of these confrontations, we determine how competitive (or fit) each strategy is. Finally, for each player, we cross over the most competitive strategies together to generate new populations of strategies. This way, features that make strategies competitive are passed on to the next generation. These new populations replace the old ones and the genetic algorithm repeats the cycle: confrontation, evaluation and generation.

In short, the GA follows **5 steps**:

1. **Creation of initial populations** of strategies for player A and player B.
2. **Confrontation between strategies.**
3. **Evaluation of strategies' fitness** based on the results of the confrontations.
4. **Generation of new populations of strategies** from the fittest strategies.
5. **Repeat from step 2.**

The following sections explore in more depth the concepts of population of strategies, fitness, confrontation, evaluation and generation of strategies. At the end of the section, we will be able to run the GA.

## Population of Strategies
A population of strategies is simply a collection of strategies for a player. For player A, it is a matrix where each column correspond to a strategy. For player B, it is a three-dimensional array where each layer (dimension `3`) corresponds to a strategy. The following code creates populations of ten random strategies for player A and player B. To easily identify the strategies, we name them `r paste("s", 1 : 10, sep="")`.
```{r setup population}
n_strategy <- 10
name_strategy <- paste("s", 1 : n_strategy, sep="")

dim_pop_A     <- c(n_card, n_strategy)            # for convenience
dimname_pop_A <- list(cards, name_strategy)       # for convenience
dim_pop_B     <- c(n_card, n_bet, n_strategy)     # for convenience
dimname_pop_B <- list(cards, bets, name_strategy) # for convenience

set.seed(123)
pop_A <- array(sample(x = bets, size = prod(dim_pop_A), replace = T),
               dim = dim_pop_A, dimnames = dimname_pop_A)
# Population of random strategies for player A
print(pop_A)
set.seed(123)
pop_B <- array(sample(x = c("Call", "Fold"), size = prod(dim_pop_B), replace = T),
               dim = dim_pop_B, dimnames = dimname_pop_B)
# Population of random strategies for player B (2 strategies)
print(pop_B[,,1:2])
```

### Cleaning Player B's Strategies
We can slightly modify player B's strategies to make them more realistic. Following common sense, we impose that player B *calls* if her/his card is the highest or if player A bets 0. Indeed, in both cases, player B risks nothing by calling player A's bet: those hands either end up in a draw or a win for player B.
```{r function clean_pop_B}
clean_pop_B <- function(pop, cards){
  pop[max(cards),    ,] <- "Call"
  pop[          , "0",] <- "Call"
  return(pop)
}

pop_B <- clean_pop_B(pop = pop_B, cards = cards)
# Clean population for player B
print(pop_B[,,1:2]) # Compare with the strategies before the cleaning (printed above).
```

## Fitness of a Strategy
The notion of *fitness* is central to the GA: it allows the algorithm to select the best performing strategies from which to generate new strategies In our case, the goal of the GA is to generate *profitable* strategies. An obvious measure of fitness of a strategy is thus how much money it yields, on average, against the strategies of the other player's population.

### Confrontations: one v. one
To determine the fitness of a strategy, we must thus determine its average gain/loss against each strategy of the other player's population. The function `confront()` helps us do that. It follows the matrix-oriented approach to simulate all possible hands between two given strategies and then return the average gain/loss of player A's strategy when opposed to that of player B. (To make `confront()` run faster, we create `win_game` outside of the function and consequently adapt how we compute `gain_A`.)

```{r function confront}
win_game <-   win_game <- array(numeric(n_card*n_card),
dim = dim_mat)
upperTriangle(win_game, diag = F) <- 1
lowerTriangle(win_game, diag = F) <- -1

confront <- function(strategy_A, strategy_B, n_card, dim_mat, bets, ante, win_game){
  
  # Matrix-oriented approach
  bet_A    <- array(rep(strategy_A, each = n_card       ) , dim = dim_mat)
  action_B <- array(strategy_B[ , match(strategy_A, bets)], dim = dim_mat)
  gain_A                   <- (ante + bet_A) * win_game
  gain_A[action_B=="Fold"] <-  ante
  
  # Average gain for player A
  return(mean(gain_A))
}
```

Here is an example of the function `confront()` in action.
```{r confront example}
strat_A <- pop_A[ , 3]
print(strat_A)

strat_B <- pop_B[ , , 1]
print(strat_B)

average_gain_A <- confront(strategy_A = strat_A, strategy_B = strat_B,
                           n_card = n_card, dim_mat = dim_mat, bets = bets,
                           ante = ante, win_game = win_game)
# Average gain of player A
print(average_gain_A)
```

This means that, on average, `strat_A`, wins `r average_gain_A` for each hand played against `strat_B`. Since the matrix-oriented approach simulates each possible hand once, it computes the *true* average gain of `strat_A` when opposed to `strat_B`.

### Confrontations: all v. all
The function `confront_populations()` uses two loops to make each strategy of player A play against each strategy of player B and return the result of these confrontations. In the loops, we use the function `confront()` to determine the average gain of player A's strategy against that of player B and record the value in the matrix `fitness` whose columns and rows respectively correspond to player A's and player B's strategies.
```{r function confront_all}
fitness <- array(NA, dim=c(n_strategy, n_strategy),
                 dimnames = list(name_strategy, name_strategy))

confront_populations <- function(pop_A, pop_B, name_strategy, fitness, n_card,
                                 dim_mat, bets, ante, win_game){
  for(strat_a in name_strategy)
    for(strat_b in name_strategy)
      fitness[strat_b, strat_a] <- confront(strategy_A = pop_A[ , strat_a  ],
                                            strategy_B = pop_B[ , , strat_b],
                                            n_card = n_card, dim_mat = dim_mat,
                                            bets = bets, ante = ante, win_game = win_game)

  return(fitness)
}
```

Here is an example of the function `confront_populations()` in action.
```{r fitness I}
print(pop_A)
print(pop_B[, , 1:2])

fitness <- confront_populations(pop_A = pop_A, pop_B = pop_B, fitness = fitness,
                                name_strategy = name_strategy, win_game = win_game,
                                n_card = n_card, dim_mat = dim_mat, bets = bets, ante = ante)
# Average gains of player A's strategies
print(fitness)
```

Taking the `colMeans` of `fitness` gives us the level of fitness player A's strategies and `rowMeans` of `-fitness` the level of fitness of player B's strategies.
```{r fitness II}
# Fitness of player A's strategy
sort(colMeans( fitness), decreasing = T)
# Fitness of player B's strategy
sort(rowMeans(-fitness), decreasing = T)
```
Let us have a closer look at these results and focus on player A' strategies (easier to analyze than player B's). `s5` is the best performing strategy against `pop_B`. It is a conservative strategy with small bets for small cards and large bets for large cards. With `s5`, player A wins on average `r colMeans(fitness)["s5"]` per hand against player B's strategies. On the opposite spectrum, `s9` and `s10` have the lowest level of fitness. Both strategies prescribe to bet `8` (large bet) for a `0` (lowest card) and `s10` even prescribes to bet `2` for a `4` (highest card)! Although the large bets for small cards could be bluffs and are thus not bad per se, betting `2` for the highest card is definitely a bad idea in this one-round version of poker. Player A loses on average `r -colMeans(fitness)["s9"]` with `s9` and `r -colMeans(fitness)["s10"]` with `s10`.

## Generating New Strategies
Now that we have the strategies' fitness, we can generate a new populations for the two players following **three steps**:

1. **Parent Selection**
2. **Crossover**
3. **Mutation**

For convenience, we use the term *parent strategy* to refer to a strategy from which new strategies are generated and *child strategy* to denote a newly generated strategy. In the following, we illustrate how to generate five child strategies for player A. The approach is fundamentally the same for player B .

### Parent Selection
We select the fittest strategies from `pop_A` to form the set of parent strategies `parents`. In our example, we select the seven (arbitrarily chosen number) fittest strategies.
```{r parent selection}
n_parents  <- 7
n_children <- 5

fitness_sorted  <- sort(colMeans(fitness), decreasing  = T)
fitness_parents <- head(fitness_sorted, n_parents)
# Name and fitness of the set of parent strategies
print(fitness_parents)

name_parents    <- names(fitness_parents)
parents         <- pop_A[ , name_parents]
# Set of parent strategies
print(parents)
```

### Children Generation

To generate child strategies, we cross over the parent strategies and introduce random mutations. In the crossover step, we randomly combine parent strategies, element by element, to generate the child strategies. In other words, the first element of a child strategy corresponds to the first element of a randomly selected parent strategy, the second element to the second element of a randomly selected parent, etc. In the mutation step, we randomly alter a small proportion of the child strategies' elements.

#### Children Generation: Crossover
We first use the function `sample()` on the parent strategies' names to create `pop_child` which represents the population of child strategies. At this stage, `pop_child`'s entries indicate the parent strategy from which each of its elements is inherited. Note that in `sample()`, we set `prob = exp(fitness_parents)` so that children strategies are more likely to inherit their elements from fitter parent strategies while ensuring that the probabilities are positive.
```{r crossover I}
set.seed(123)
pop_child <- array(sample(name_parents, size = n_card*n_children, replace = T,
                         prob = exp(fitness_parents)), # exp() ensures positive probabilities
                  dim = c(n_card, n_children),
                  dimnames = list(cards, paste("Child", 1 : n_children)))
# Origin of the element of the child strategies
print(pop_child)
```
The first and the third elements of `Child 1` come from the parent strategy `s8`, the second element from `s2`, etc.

Next, we loop through the the names of the parent strategies. In the loop, we first assign to `strategy_parent` the parent strategy with the appropriate name. We then assign to `location_parent` the location of the elements of `pop_child` that are inherited from `strategy_parent`. Finally, we substitute these elements inherited from `strategy_parent` with the corresponding elements of `strategy_parent`. (Since the dimensions of `pop_child` and `strategy_parent` are different, we use `rep()` on the latter.)
```{r crossover II}
for(name_parent in name_parents){
  
  strategy_parent <- parents[ , name_parent]
  location_parent <- pop_child == name_parent
  
  pop_child[location_parent] <- rep(strategy_parent, n_children)[location_parent]

}

pop_child <- array(as.numeric(pop_child),
                   dim = c(n_card, n_children),
                   dimnames = list(cards, paste("Child", 1:n_children)))
# Population of children strategies
print(pop_child)
```

#### Children Generation: Mutations
Finally, we introduce random mutations to `pop_child`. We first assigne the desired mutation rate to `mutation_rate` and use the function `sample()` with `prob = c(mutation_rate, 1 - mutation_rate)` to create the matrix `mutation_location` which indicates the random location of the mutations. We then use the function `sample()` again to generate the vector `mutation_outcome` which indicates the random outcome of the mutations. Finally, we subsitute the elements of `pop_child` where a mutation occurs with the values of `mutation_outcome`.

```{r mutation}
mutation_rate <- 0.2

mutation_location <- array(sample(c(T, F), size = n_card * n_children, replace = T, 
                                  prob = c(mutation_rate, 1 - mutation_rate)),
                  dim = c(n_card, n_children), 
                  dimnames = list(cards, paste("Child", 1 : n_children)))
# Location of mutations
print(mutation_location)

n_mutation       <- sum(mutation_location)
mutation_outcome <- sample(bets, size = n_mutation, T)
# Outcome of mutations
print(mutation_outcome)

pop_child[mutation_location] <- mutation_outcome
# Population of child strategies after mutations
print(pop_child)
```

### Code
The functions `generat_A` and `generate_B` encapsulate these three steps for player A and player B. These functions are the last one we need to be able to run the GA.

```{r function generate_A}
generate_A <- function(fitness, pop, n_parents, dim_pop_A, dimname_pop_A,
                       n_strategy, bets, mutation_rate){
  
  # Parent Selection
  fitness_strategy <- colMeans(fitness)
  fitness_parents  <- head(sort(fitness_strategy, decreasing  = T), n_parents)
  name_parents     <- names(fitness_parents)
  parents          <- pop[ , name_parents]
  
  # Crossover
  pop <- array(sample(name_parents, size = prod(dim_pop_A), replace = T,
                      prob = exp(fitness_parents)),
               dim = dim_pop_A, dimnames = dimname_pop_A)
  
  for(parent in name_parents){
    
    strategy_parent <- parents[ , parent]
    location_parent <- pop == parent
    pop[location_parent] <- rep(strategy_parent, n_strategy)[location_parent]

  }
  
  pop <- array(as.numeric(pop), dim = dim_pop_A, dimnames = dimname_pop_A)
  
  # Mutation
  mutation_location <- array(sample(c(T, F), size = prod(dim_pop_A), T,
                                    prob = c(mutation_rate, 1-mutation_rate)),
                             dim = dim_pop_A, dimnames = dimname_pop_A)
  
  n_mutations      <- sum(mutation_location)
  mutation_outcome <- sample(bets, size = n_mutations, replace = T)
  
  pop[mutation_location] <- mutation_outcome
  
  return(pop)
  
}
```

```{r function generate_B}
generate_B <- function(fitness, pop, n_parents, dim_pop_B, dimname_pop_B,
                       n_strategy, mutation_rate, cards){
  
  # Parent Selection
  fitness          <- - fitness
  fitness_strategy <- rowMeans(fitness)
  fitness_parents  <- head(sort(fitness_strategy, decreasing  = T), n_parents)
  name_parents     <- names(fitness_parents)
  parents          <- pop[ , , name_parents]
  
  # Crossover
  pop <- array(sample(name_parents, size = prod(dim_pop_B), replace = T,
                      prob = exp(fitness_parents)),
               dim = dim_pop_B, dimnames = dimname_pop_B)
    
  for(parent in name_parents){
    
    strategy_parent <- parents[ , , parent]
    location_parent <- pop == parent
    
    pop[location_parent] <- rep(strategy_parent, n_strategy)[location_parent]
    
  }
  
  # Mutation
  mutation_rate <- 2 * mutation_rate
  # since player B has only two actions i.e. "Call" or "Fold"
  # half of the mutations have no effect.
  mutation_location <- array(sample(c(T, F), size = prod(dim_pop_B), replace = T,
                                    prob = c(mutation_rate, 1-mutation_rate)),
                           dim = dim_pop_B, dimnames = dimname_pop_B)
  
  n_mutations            <- sum(mutation_location)
  mutation_outcome       <- sample(c("Call", "Fold"), n_mutations, T)
  pop[mutation_location] <- mutation_outcome
  
  pop <- clean_pop_B(pop = pop, cards = cards)
  
  return(pop)
  
}
```

Here is an example of the function `generate_A` in action.
```{r generate_A example}
pop_A_child <- generate_A(fitness = fitness, pop = pop_A, n_parents = n_parents,
                         dim_pop_A = dim_pop_A, dimname_pop_A = dimname_pop_A,
                         n_strategy = n_strategy, bets = bets, mutation_rate = 0.05)

# Parent Strategies
print(pop_A)       # random bets for small cards
# Child Strategies
print(pop_A_child) # small bets for small cards.
```

# The Genetic Algorithm in Action
Now that we have covered all the components of the GA, we can finally write the function `my_GA` which encapsulates them.

```{r function my_GA}
my_GA <- function(cards = 1:10, bets = seq(0,20,2), ante = 5, n_strategy = 200,
                  n_generations = 10, prop_parents = 2/3, mutation_rate = 0.05){
  
  # Setup
  n_card        <- length(cards)
  n_bet         <- length(bets)
  n_parents     <- n_strategy * prop_parents
  
  name_strategy <- paste("s", 1 : n_strategy, sep = "")
  
  dim_pop_A     <- c(n_card, n_strategy)
  dim_pop_B     <- c(n_card, n_bet, n_strategy)
  dim_mat       <- c(n_card, n_card)
  dim_fit       <- c(n_strategy, n_strategy)
  dim_output_A  <- c(dim_pop_A, n_generations)
  dim_output_B  <- c(dim_pop_B, n_generations)
  dim_fit_A     <- c(dim_fit, n_generations)
  
  dimname_pop_A    <- list(cards, name_strategy)
  dimname_pop_B    <- list(cards, bets, name_strategy)
  dimname_mat      <- list(cards, cards)
  dimname_fit      <- list(name_strategy, name_strategy)
  dimname_output_A <- list(cards, name_strategy, 1 : n_generations)
  dimname_output_B <- list(cards, bets, name_strategy, 1 : n_generations)
  dimname_fit_A    <- list(name_strategy, name_strategy, 1 : n_generations)
  
  win_game      <- array(numeric(n_card * n_card), dim = dim_mat)
  upperTriangle(win_game, diag = F) <-  1
  lowerTriangle(win_game, diag = F) <- -1
  
  fitness <- array(NA, dim = dim_fit, dimnames = dimname_fit)
  
  output_A  <- array(NA, dim = dim_output_A, dimnames = dimname_output_A)
  output_B  <- array(NA, dim = dim_output_B, dimnames = dimname_output_B)
  fitness_A <- array(NA, dim = dim_fit_A   , dimnames = dimname_fit_A   )

  # Initialization
  generation <- 1
  
  pop_A <- array(sample(bets, size = prod(dim_pop_A), replace = T),
                 dim = dim_pop_A, dimnames = dimname_pop_A)
  pop_B <- array(sample(c("Call", "Fold"), size = prod(dim_pop_B), replace = T),
                 dim = dim_pop_B, dimnames = dimname_pop_B)
  pop_B <- clean_pop_B(pop_B, cards = cards)
  
  output_A[ , , generation  ] <- pop_A
  output_B[ , , , generation] <- pop_B
  
  
  # Loop
  while(generation < n_generations){

    fitness <- confront_populations(pop_A = pop_A, pop_B = pop_B,
                                    name_strategy = name_strategy,
                                    fitness = fitness, n_card = n_card,
                                    dim_mat = dim_mat, bets = bets, 
                                    ante = ante, win_game = win_game)

    pop_A <- generate_A(fitness = fitness, pop = pop_A, n_parents = n_parents,
                        dim_pop_A = dim_pop_A, dimname_pop_A = dimname_pop_A,
                        n_strategy = n_strategy, bets = bets,
                        mutation_rate = mutation_rate)
    
    pop_B <- generate_B(fitness = fitness, pop = pop_B, n_parents = n_parents,
                        dim_pop_B = dim_pop_B, dimname_pop_B = dimname_pop_B,
                        n_strategy = n_strategy, cards = cards,
                        mutation_rate = mutation_rate)
    
    # Output
    fitness_A[ , , generation ] <- fitness
    
    generation <- generation + 1
    output_A[ , , generation  ] <- pop_A
    output_B[ , , , generation] <- pop_B
    
  } # close for-loop
  
  return(list(A = output_A, B = output_B,
              fit = fitness_A))
  
}
```

Finally, the GA in action:
```{r my_GA in action, cache = TRUE}
set.seed(123)
results <- my_GA()

results_A <- results$A
results_B <- results$B

# 10 strategies of the 5th generation for player A
results_A[ , 1 : 10, 5]

# 2 strategies of the 10th generation for player B
results_B[ , , 25 : 26, 10]
```

```{r, echo = FALSE}
save(clean_pop_B,
     confront, confront_populations,
     generate_A, generate_B,
     my_GA,
     file = "functions.RDATA")
```
