---
title: "Results"
author: "Raphaël Morsomme"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    number_sections: TRUE
    fig_height: 12
    fig_width: 10
runtime: shiny
---

# Introduction
We experiment with the parameters of our GA to anwser two questions: (i) **what conditions encourage player A to bluff** and (ii) **is there any trick that player B can adopt to stop player A's bluff**. It appears that the size of the ante and the mutation rate have an impact on player A's level of bluff. Large antes make a successful bluff more profitable and are associated with larger levels of bluff for player A. Low mutation rates make the environment more stable and are also associated with larger levels of bluff. After discussing these results, I mention ways to improve the current analysis.

## Defining Bluff
Since our version of poker only allows player A to bluff, we exclusively focus on player A's strategy and how often (s)he bluffs.
We use a simple definition of bluff: *player A bluffs if (s)he bets large amounts for small cards*. In this script, we assess whether player A bluffs *visually*. The following two plots display two strategies for player A, one with, and one without bluff. They should guide our interpretation of the results.

```{r, fig.height = 5, echo = F}
par(mfrow=c(1,2), cex = 1.1)

barplot(c(12, 7, 8, 2, 3, 3, 12, 15, 18, 17),
        names.arg = 1:10,
        ylim = c(0, 20),
        main = "Strategy for Player A with Bluff",
        xlab = "Player A's Card", ylab = "Bet")
abline(h=seq(0,20,2), lty=2)

barplot(c(2, 1, 2, 2, 3, 3, 12, 15, 18, 17),
        names.arg = 1:10,
        ylim = c(0, 20),
        main = "Strategy for Player A without Bluff",
        xlab = "Player A's Card", ylab = "Bet")
abline(h=seq(0,20,2), lty=2)

par(mfrow=c(1,1), cex = 1)
```

# Results
Two factors influence player A's level of bluf: the size of the ante and the mutation rate. We run the GA with three levels of ante (`0`, `5` and `10`) and four levels of mutation rate (`0.001`, `0.01`, `0.05` and `0.1`).

```{r import results, echo = FALSE}
load("results.RDATA")
```

```{r, echo = FALSE}
radioButtons("dataset", "Parameters",
             choiceNames  = list("default (ante = 5, mutation rate = 0.05)",
                                 "ante = 0", "ante = 10",
                                 "mutation rate = 0.001", "mutation rate = 0.01", "mutation rate = 0.1"),
             choiceValues = list("results_ante_05_mut_05",
                                 "results_ante_00_mut_01", "results_ante_10_mut_05",
                                 "results_ante_05_mut_001", "results_ante_05_mut_01", "results_ante_05_mut_10")
             )

sliderInput("generation", "Generation", value = 1, min = 1, max = 125, step = 1)

renderPlot({
  
  par(mfrow = c(1, 2), cex = 1.1)
  dataset <- get(input$dataset)
  
  barplot(rowMeans(dataset$A[ , , input$generation]), 
          ylim = c(0, 20),
          main = "Average Strategy for Player A",
          xlab = "Player A's Card",
          ylab = "Average Bet")
  abline(h=seq(0,20,2), lty=2)
  
  image(1:10, seq(0,20,2),
        rowMeans(dataset$B[ , , , input$generation] == "Call", dims = 2),
        col = gray((100 : 0) / 100),
        main = "Average Strategy for Player B",
        xlab = "Player B's Card", ylab = "Player A's bet")
  })

renderText({
  
  dataset <- get(input$dataset)
  
  print(paste("Player A bets on average", round(mean(dataset$A[ , , input$generation] ), 2), "per hand."))
  
  })

renderText({
  
  dataset <- get(input$dataset)
  
  print(paste("Player A gains on average", round(mean(dataset$fit[ , , input$generation] ), 2), "per hand."))
  
  })
renderText({
  
  dataset <- get(input$dataset)
  
  print(paste("Player B calls ", round(100 * mean(dataset$B[ , , , input$generation]  == "Call"), 1), "% of the time.", sep = ""))
  
  })

renderPlot({
  
  par(mfrow = c(1, 2), cex = 1.1)
  dataset <- get(input$dataset)
  fitness <- dataset$fit[ , , input$generation]
  strat_fittest_A <- names(sort(colMeans(fitness), decreasing  = T)[1])
  strat_fittest_B <- names(sort(rowMeans(-fitness), decreasing  = T)[1])
  
  barplot(dataset$A[ , strat_fittest_B, input$generation], 
          ylim = c(0, 20),
          main = "Best Strategy for Player A",
          xlab = "Player A's Card",
          ylab = "Bet")
  abline(h=seq(0,20,2), lty=2)
  
  image(1:10, seq(0,20,2),
        dataset$B[ , , strat_fittest_A, input$generation] == "Call",
        col = gray((100 : 0) / 100),
        main = "Best Strategy for Player B",
        xlab = "Player B's Card",
        ylab = "Player A's bet")
  })
```

## Ante
The size of the ante determines how profitable it is for player A to bluff. Indeed, player A wins the ante if player B does not call the bet. Large antes encourage player A to bluff and small ones have the opposite effect. In the extreme case, when `ante = 0`, player A should never bluff as (s)he can only lose by doing so: even if player B does not call the bet, (s)he wins nothing (ante is `0`).

This is exactly what we find. When the ante is `0`, player A adopts an extremely conservative strategy (see generation `25` and after). After some time (generation `93` and after), player A's strategies become extreme and (s)he only bets for the card `10`. Note that I have reduced the mutation rate from `0.05` to `0.01` to accentuate this pattern. On the contrary, when we increase the ante to `10`, player A bluffs much more often (e.g. best strategy of generation `118` with bets of `20` -- bluffs -- for the two lowest cards.).

This shows that, unsurprisingly, the more profitable the bluff, the more player A bluffs. This has the following consequence for situations involving negotiation: one should bluff (and expect other people to bluff) if, and only if, it is worth it. This means that a way to prevent someone else form bluffing is to convince her/him that bluffing is not worthwhile, that the gain of a bluff are too small for the bluff to be profitable. As for poker, this means that a way to encourage players to bluff -- and thereby make the games more exciting -- is simply to increase the value of the ante.

## Mutation Rate
The link between stability, and bluff is not obvious at first glance (at least it was not to me). Yet, we observe that when the mutation rate is low (`0.01` and `0.001`), player A bluffs more often and player B adopts a cautious strategy. Note that if the mutation rate is too small (`0.001`), then the population become homogeneous and evolve in an erratic fashion. When the mutation rate is large (`0.1`), then player A bluffs adopts a more conservative strategy and player B is more adventutous.

This can be explained by the fact that in our case, low levels of mutation rate result in cautious strategies for player B where (s)he folds for large bets of player A. The child generations of player B closely copy (low mutation rate) the most profitable parent strategies of the previous generation which happened to fold for large bets of player A (see `mutation rate = 0.01`, generation 37 where player B folds for large bets of player A). The adoption of cautious strategies by player B in turn encourages player A to bluff since her/his bluffs will not be called. Indeed, bluff has the best chance to succeed when the opponent is cautious. On the other hand, bluff has no chance to succeed when the opponent is adventurous (the opposite of cautious). In our case, this occurs when the child generations of player B depart (large mutation rate) from their parents, calling, as a result, player A's bluffs more often.

This implies, that in very stable situations where people are cautious, bluffing is most profitable. This means that a way to protect oneself from the bluffs of others, is to be more adventurous and call the their bluffs more often.

## Other Parameters
I also explored the effect of the other parameters of the GA on player A's level of bluff and found that they do not influence it. Increasing the number of cards or bets only slows down the algorith and makes the patterns need a larger number of generations to appear. The number of strategies in the population also has no impact on player A's level of bluff. Yet, one needs a sufficient amount of strategies in the populations for the algorithm to work. If there are not enough strategies in the population (`20` for instance), then the strategies struggle to evolve after a few generations. Finally, the proportion of strategies that we select to form the set of parent strategies has to be, to my surprise, relatively large. A small value also results in population that struggle to evolve.

# Looking Forward
The current analysis could be improved in several ways.

## Operationalization
So far, we have conducted a visual assessment of player A's level of bluff. Yet, it would be beneficial to measure it numerically and therey be able to conduct statistical test on the results. We could for instance measure the distance between player A's strategies and a typically conservative strategy.

Similarly, we could also come up with a measure of how cautious player B's strategies are. Again, we could measure the distance between player B's strategies and some baseline strategy.

## Elitism
Note that our version of the GA does not include *elitism*. Elitism refers to the practice of preserving the best parent strategies (the elites) in the following generation. If we set the proportion of elites to, say, `10%`, then the `10%` fittest strategies are kept and the children strategies replace the remaining strategies. Elitism ensures that, regardless of the mutations occuring in the children strategies, we still have good strategies (the elites) in the new population, thereby making the GA more efficient.

## Stochastic Strategies
Finally, we could introduce a *stochastic* element in the strategies of the players. For player A, this could take the form of a strategy that gives a density distribution for each card (s)he could receive. Player A's bet then corresponds to a value randomly chosen according to the density distribution. Such strategy could for instance consists of matrix with two columns which provide a mean and a standard deviation for a normal distribution for each possible card. The current version of player A's strategy is a special case of such stochastic strategy where the standard deviations are `0`. As for player B, the strategy could indicate the probability of calling player A's bet given her/his card and player A's bet. The current version of player B's strategy is a special case of such stochastic strategy where the two allowed probabilities are `0` (fold) and `1` (call). Designing strategies containing a stochastic element will be the topic of my next script.